On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2023-09-15 Manuscript/manuscript.tex
Fixed ordering of glmm/sign rank and went over Prof Schifano's notes

%Abstract should be 200 words
%Devon Allen was disqualified in 2022... discussions about dq, talk methods and 
%models, venue effect, and conclusion
\begin{abstract}
Devon Allen was disqualified in the men's 110 meter hurdle final of the 2022
World Track and Field Championships after registering a reaction time of 0.099 
seconds, 0.001 seconds faster than what is allowed. Following the games, 
bloggers on the running website LetsRun concluded that the reaction times  
from the 2022 World Championships seemed to be generally faster compared to the 
other reaction times they considered, but they did not perform any formal 
statistical analysis. This paper questions the reaction time 
disqualification barrier, which is currently 0.1 seconds, to determine whether 
this is a reasonable threshold. We employ a generalized linear mixed model 
(GLMM) with random venue (year) effect and a random heat effect within venue 
in order to model the reaction time data. 
Additionally, we employ a signed-rank test for clustered data to 
compare reaction times for the same athletes at different competitions. This 
matter needs to be addressed because disqualification based on allowable 
reaction time will continue to be an issue in future world championships.

@@

Devon Allen, an alumnus of the University of Oregon in Eugene, Oregon, was
expected to be a contender for the 2022 World Track and Field Championships 110 
meter hurdle.  After running the event in 12.84 seconds in early June 2022 
(only 0.04 seconds away from the world record \citep{wa2022preview}) and coming 
in third at the United States Track and Field Championships also in Eugene 
in late June, Allen was poised to compete in the 2022 World Track and 
Field Championships in August, which would also be held in Eugene, Oregon.
At the World Championships, Allen made it through the preliminary
heats and semifinals to reach the final heat while posting reaction times of 
0.123 and 0.101 in his preliminary heat and semifinal, respectively.  
His disqualification in the final heat by reacting in 0.099 seconds, just 0.001
seconds faster than the 0.1 threshold set by the International Association of
Athletics Federations (IAAF), was met
with shock from both Allen and fans alike.  NBC, who was broadcasting the
championships in the United States, later uploaded a video to their YouTube
channel that showed the disqualification
(\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}).
The crowd erupted with jeers
when they realized that Allen was disqualified, angry at the result.  Had Allen 
reacted just 0.001 seconds slower he would have been allowed to compete. %and been 
%spared waiting until the next World Track and Field Championships.
%\eds{He is not competing this year, correct? If yes, we may need to rephrase.}
%\of{Correct}
%\eds{removed because he did not compete this year, so did not actually have
% to wait, after all.} 
\eds{ If it is true he ``retired'' as a runner to play football, we could 
instead say something like
\textit{Had Allen reacted just 0.001 seconds slower he would have been allowed 
to compete before retiring from running to play football.}}

@@

The objective to this paper is to analyze the reaction times at the 2022 World 
Track and Field Championships to check for any abnormalities, and we approach
this in two different ways.  First we model the data, which are
the reaction times of every non-disqualification since the 1999 World 
Championships, using a gamma linear mixed-effects model with two random effects: 
one effect for the year and a second effect for the
heat within each year.
Based on the fitted model, we can then estimate the probability of a reaction 
time being below a certain threshold (e.g., 0.1). 
Second, we compare reaction times for athletes who have competed
at additional competitions to the 2022 World Track and Field Championship.  By
comparing reaction times from the same athlete across multiple competitions we 
are able to determine if the 2022 reaction times differed significantly from 
those of other competitions.
Our investigation into reaction times is important not only because of its
implications for future races, but also because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times, which is the goal of
every sprinter who steps onto the track.
%\eds{Since the World Championships are happening again now, is data from this 
%year already available? If it is available, we would need to justify its omission
%in the current analysis.
%If it is not available, but becomes available while under review, would we 
%reanalyze the data?  Does it seem likely our conclusions would change based on
%what is going on this year?}
%\of{Resolved}

The rest of this paper is organized as follows. Section~\ref{sec:Data} describes 
how the data were collected and begins to detail the generalized linear mixed-effect 
 model (GLMM) that is further developed in Section~\ref{sec:glmm}. The 
signed-rank test that analyzes reaction times for the same athletes across
different competitions is
also described in Section~\ref{sec:rank}.  Section~\ref{sec:Results}
 provides the results of the two analyses and ends 
by questioning the reaction time barrier.  Lastly, Section~\ref{sec:concludingremarks}
discusses the impact and limitations of the paper.

@@

\subsubsection{2022 National Competitions}\label{sec:datanational}
From June 23-26, 2022, the United States held its Track and Field Championships 
at Hayward Field in Eugene, Oregon to decide who to send to the World 
Championships being held at the same venue in August. Thus, we can establish a 
baseline for the four United States athletes who competed in at least one round 
of both events: Trey Cunningham, Daniel Roberts, Grant Holloway, and Devon Allen.   
Every athlete reacted faster in all the World Track and Field Championships 
races compared to the United States Track and Field Championships races. The 
dataset of the four United States athletes is too small to perform any 
meaningful statistical analysis, but the data was expanded to include reaction 
times for 11 more athletes who competed at other national competitions (many of which were 
held from May-July of 2022). This data was difficult to find, as the data is not
centrally located but instead found on different national websites (often 
recorded in the national language without translation). 
\of{I don't know how to say this}. \eds{is my change correct?} \of{No I was able
to use a lot even though I couldn't actually read the page. I just tried to find
110 meter hurdle and then look at a column for reaction times which are always from
roughly 0.1 to 0.18 seconds.} \eds{is it correct now?}

@@

While it is theoretically possible that athletes got faster 
between 2019 and 2022 or 2022 and 2023, it is important to 
remember that these are some of the fastest sprinters on the planet, and thus
improving on elite times is not easy.  Once again we repeat a similar procedure
as in Section~\ref{sec:datanational}; we assign each athlete to be a cluster and
within each cluster label times to be from either the 2019, 2022, or 2023 World
Championships. When comparing 2019 and 2022 we look at data from 15
athletes, with two clusters of size two, three clusters of size three, five 
clusters of size four, two clusters of size five, and three clusters of size six.
When comparing 2022 and 2023 we look at data from 22 athletes, with seven 
clusters of size two, three clusters of size three, six clusters of size four,
and six clusters of size five.  Although 2022 is presented three times in
Figure~\ref{fig:RankScatterplots}, the subset of athletes differs each time.
There may be athletes who competed in 2019 and 2022 and had their national
championships data available, but they did not compete in 2023, and thus their
times appear only in the top two graphs. \eds{top and bottom graphs?}
\eds{also, you will likely need to make Fig 1 smaller as the caption runs off
the page (even before my edits).}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{RankScatterPlots}
  \caption{The top graph shows the reaction times of athletes who competed at 
	least once each at the 2019 and 2022 World Championships. The middle graph 
	shows the reaction times of athletes who competed at least once each at the 
	2022 and 2023 World Championships. The bottom graph shows the reaction times 
	of athletes who competed at least once at their country's
  2022 national championships and the 2022 World Championships.
	\eds{check x-axis labels in the bottom plot: last three should be 2022?}}
  \label{fig:RankScatterplots}
\end{figure}

Figure~\ref{fig:RankScatterplots} shows the data that will be analyzed
with the rank-based methods (described in Section~\ref{sec:rank} with results in 
Section~\ref{subsec:Results_Rank}). In each of the three graphs, the reaction 
times for athletes who competed
in the 2022 finals \eds{World Championships?} are compared to their reaction
times at another at which they competed.  
It is worth noting that each instance of the 2022 data is different, because
those who competed at both the 2019 and 2022 World Championships will be
different athletes than those who competed at both the 2022 and 2023 World
Championships. This data was only for athletes who competed
in at least one race at each championship, however it is possible for athletes
to have competed in one race in 2019 and two races in 2022. 
\eds{did we not already cover the last two sentences in the previous paragraph, 
or are you trying to say something different here?}  The 2022 World
Championship Heats and Semifinals both had on average lower reaction times than
their respective 2019 races, and the 2022 finals were substantially faster. It
is also worth noting is that the fastest reaction time in the 2022 World
Championships Finals and Semifinals both came from Devon Allen, but the 
difference between 0.101 and 0.099
seconds was the difference in what caused him to be disqualified. 
%\eds{Please reorganize the labels so they are either in one of the following 
%orders:
%2019 Heat, 2019 Semi, 2019 final, 2022 Heat, 2022 Semi, 2022 final  OR
%2019 Heat, 2022 Heat, 2019 Semi, 2022 Semi, 2019 final, 2022 final.  
%(I am not sure which is better, but perhaps the latter?)}
%\of{I re-worked this graph completely, please let me know what you think.  For
%each graph I currently list them in chronological order: so 2022 then 2023 but
%I list 2022 second when comparing it to 2019/rank comparison.  Please let me
%know your thoughts on this and if I should list 2022 and then 2023 for visual
%consistency.}
%\eds{I think the graphs look good}

We also started to perform a similar analysis to compare the times of female 
athletes who competed at the 2019 and 2022 World Track and Field Championships,
however as our results were not significant, we decided to incorporate another
method to investigate the times at the 2022 World Championships and thus we
employ the GLMM, which looks at historical men's data to try to determine
whether Allen's time and disqualification are reasonable. 
\eds{I would suggest removing this paragraph, particularly if you still include
the comments about female reaction times in the Concluding Remarks. 
No analysis has been done yet, nor have the methods been described yet, so it 
does not make sense in my opinion 
to refer to significance or another analysis method here.}

@@

Data was taken from the World Athletics 
and covers the men's 110 meter hurdles from 1999 to 2023; the source can be 
found in the Appendix. 
Although women's data is available, we chose not to include women's data in the
analyses to follow due to the numerous papers citing differences in men and
women's reaction times \citep[e.g.,][]{lipps2011implications, 
babicc2009reaction, panoutsakopoulos2020gender}.
\eds{you should mention the data source earlier, possible in Section 2 
before Section 2.1 if it applies to Section 2.1.1, and also the sources of the 
national data for 2.1.2 (i.e., pointing to the info in Appendix).}
For the creation of the GLMM model we focus on the reaction times during the 
semifinal and final heats only, as reaction times from preliminary heats are 
often not as fast as those in later heats \citep[e.g.,][]{zhang2021correlation}.

@@

With our small sample size, the asymptotic normal distribution may not be
reliable, so we use 1000 random permutations to simulate the observed test 
statistic.  \eds{simulate the null distribution of the test statistic?}
%\eds{I do not understand what is meant in the text after the colon in the 
%previous sentence.  Please rephrase.}
%\of{I'm not sure what that was, I removed it now.}
This method is available from the \texttt{clusWilcox.test()} function
with \texttt{method = `ds'} (for \underline{D}atta and \underline{S}atten) and 
\texttt{exact = TRUE} from R package
\texttt{clusrank} \citep{jiang2020wilcoxon}.

@@

Table~\ref{tab:Clusrankresults} shows the results generated from four \eds{three?}
tests in R.
We use a \texttt{clusWilcox.test()} with \texttt{exact = TRUE} and 
\texttt{B = 1000} indicating that the random permutation size was chosen to be 
1000, as discussed previously in Section~\ref{sec:rank}.  
%As the sample sizes are small, 65 and 49, we use \texttt{exact = TRUE} as
%that is preferred.  
%In both cases, 
The national versus international comparison
produced lower p values than the 2019 versus 2022 comparison, with the
national vs international \texttt{exact = TRUE} test producing a p value of 
$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  This
is a highly significant test result and shows that athletes who competed at the
2022 World Track and Field Championships were on average, significantly faster
at reacting than they were a few months prior.  For the 2019 versus 2022
same athlete comparison with \texttt{exact = TRUE} argument, the p value was
significant at $\alpha = 0.01$, however when \texttt{exact = FALSE}, the p
value was higher and only significant at $\alpha = 0.05$. 
\eds{not sure the \texttt{exact = FALSE} results need to be reported here or in
Table 4 when we say we use \texttt{exact = TRUE} here and in the previous section.}
Nonetheless, these
tests help to show that athletes competing at the 2022 World Track and Field
Championships had faster reaction times than when they had competed previously.
\eds{need to comment on 2022 vs 2023 results}

@@

\begin{table}
  \centering
  \caption{Estimated parameters for the two effect GLMM on various data sets. SE
  is short for standard error. \eds{data in table needs to be revisited and table
	needs to be referenced within the text. Also, it needs to be clarified 
	somewhere that ``Old Pooled'' excludes 2022 and ``All Pooled'' includes 2022.}} 
  \begin{tabular}{c c c c c c}
      \toprule
      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $\delta_{i/j}$ \\
      \midrule
      Year Effect Only & $-2023.5$ & N/A & 0.048 & $-1.857$ & 0.152\\
      Heat Effect Only & $-2070.7$ & 0.073 & N/A & $-1.890$ & 0.135\\
      Full model       & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
      \bottomrule
  \end{tabular}
  \label{tab:Gamma_parameters}
\end{table}

@@

This paper does not consider data from the women's 100 meter hurdle, and although
that was initially because of concerns over reaction time differences in men and
women, exploratory analysis showed that 2022 was at best only
slightly faster relative to other championships going back to 2003 
%\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
%the intro, unless he only considered sprint and not hurdles?} \of{Johnson's
%results considered the mens 100m, mens 110m hurdles, womens 100m but not the 
%womens 100m hurdles.} 
Likewise, the rank-based comparisons resulted in p values
that were not significant. 
\eds{could that be because of even smaller sample sizes?}
  The lack of evidence from the women's data makes it 
harder to conclude that the sensors used at the 2022 championships were faulty, 
although it is unknown if the sensors for the men's 110 meter hurdle and women's
100 meter hurdle were the same.  
\eds{is any calibration of the sensors needed (e.g. for weight?) that might
also play a role for women vs men. (My husband asked me about this as a possible 
reason for the difference in results for men and women, but he did not know
if any calibration is actually done.)}
\of{I honestly am not sure, do you think that this is worth investigating, or
beyond the scope of this paper? I am not sure if the mechanics of Seiko's
technology is public knowledge.}

@@

\eds{what happened to the Appendix?}

>>>>

%Abstract should be 200 words
%Devon Allen was disqualified in 2022... discussions about dq, talk methods and 
%models, venue effect, and conclusion
\begin{abstract}
Devon Allen was disqualified in the men's 110 meter hurdle final of the 2022
World Track and Field Championships after registering a reaction time of 0.099 
seconds, 0.001 seconds faster than what is allowed. Following the games, 
bloggers on the running website LetsRun concluded that the reaction times  
from the 2022 World Championships seemed to be generally faster compared to the 
other reaction times they considered, but they did not perform any formal 
statistical analysis. This paper questions if athletes who competed at the 2022
World Track and Field Championship had similar times at other competitions and
if 0.1 seconds is a reasonable disqualification barrier. To do so, we employ a 
signed-rank test for clustered data to compare reaction times for the same 
athletes at different competitions and a generalized linear mixed model 
(GLMM) with random venue (year) effect and a random heat effect within venue 
in order to model the reaction time data. This matter needs to be addressed 
because disqualification based on allowable reaction time will continue to be an
issue in future world championships.

@@

Devon Allen, an alumnus of the University of Oregon in Eugene, Oregon, was
expected to be a contender for the 2022 World Track and Field Championships 110 
meter hurdle.  After running the event in 12.84 seconds in early June 2022 
(only 0.04 seconds away from the world record \citep{wa2022preview}) and coming 
in third at the United States Track and Field Championships also in Eugene 
in late June, Allen was poised to compete in the 2022 World Track and 
Field Championships in August, which would also be held in Eugene, Oregon.
At the World Championships, Allen made it through the preliminary
heats and semifinals to reach the final heat while posting reaction times of 
0.123 and 0.101 in his preliminary heat and semifinal, respectively.  
His disqualification in the final heat by reacting in 0.099 seconds, just 0.001
seconds faster than the 0.1 threshold set by the International Association of
Athletics Federations (IAAF), was met
with shock from both Allen and fans alike.  NBC, who was broadcasting the
championships in the United States, later uploaded a video to their YouTube
channel that showed the disqualification
(\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}).
The crowd erupted with jeers
when they realized that Allen was disqualified, angry at the result.  Had Allen 
reacted just 0.001 seconds slower he would have been allowed to compete. %and been 
%spared waiting until the next World Track and Field Championships.
%\eds{He is not competing this year, correct? If yes, we may need to rephrase.}
%\of{Correct}
%\eds{removed because he did not compete this year, so did not actually have
% to wait, after all.} 
\eds{ If it is true he ``retired'' as a runner to play football, we could 
instead say something like
\textit{Had Allen reacted just 0.001 seconds slower he would have been allowed 
to compete before retiring from running to play football.}}
\of{I don't think he retired, he was competing in races as recently as this past
June.}

@@

The objective to this paper is to analyze the reaction times at the 2022 World 
Track and Field Championships to check for any abnormalities, and we approach
this in two different ways.  First we compare reaction times for athletes who 
have competed at additional competitions to the 2022 World Track and Field 
Championship.  By comparing reaction times from the same athlete across multiple
competitions we are able to determine if the 2022 reaction times differed 
significantly from those of other competitions.  Then we model the data, which are
the reaction times of every non-disqualification since the 1999 World 
Championships, using a gamma linear mixed-effects model with two random effects: 
one effect for the year and a second effect for the heat within each year.
Based on the fitted model, we can then estimate the probability of a reaction 
time being below a certain threshold (e.g., 0.1). 
Our investigation into reaction times is important not only because of its
implications for future races, but also because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times, which is the goal of
every sprinter who steps onto the track.
%\eds{Since the World Championships are happening again now, is data from this 
%year already available? If it is available, we would need to justify its omission
%in the current analysis.
%If it is not available, but becomes available while under review, would we 
%reanalyze the data?  Does it seem likely our conclusions would change based on
%what is going on this year?}
%\of{Resolved}

The rest of this paper is organized as follows. Section~\ref{sec:Data} details 
how the two types of data were collected, as the signed-rank test as described
in Section~\ref{sec:rank} uses slightly different data than the generalized 
linear mixed-effect model (GLMM) that is further developed in 
Section~\ref{sec:glmm} Section~\ref{sec:Results} provides the results of the two
analyses and ends by questioning the reaction time barrier.  Lastly, 
Section~\ref{sec:concludingremarks} discusses the impact and limitations of the
paper.

@@

\subsubsection{2022 National Competitions}\label{sec:datanational}
From June 23-26, 2022, the United States held its Track and Field Championships 
at Hayward Field in Eugene, Oregon to decide who to send to the World 
Championships being held at the same venue in August. Thus, we can establish a 
baseline for the four United States athletes who competed in at least one round 
of both events: Trey Cunningham, Daniel Roberts, Grant Holloway, and Devon Allen.   
Every athlete reacted faster in all the World Track and Field Championships 
races compared to the United States Track and Field Championships races. The 
dataset of the four United States athletes is too small to perform any 
meaningful statistical analysis, but the data was expanded to include reaction 
times for 11 more athletes who competed at other national competitions (many of which were 
held from May-July of 2022). This data was difficult to find, as the data is not
centrally located but instead found on different national websites (often 
recorded in the national language without translation). 
\of{I don't know how to say this}. \eds{is my change correct?} \of{No I was able
to use a lot even though I couldn't actually read the page. I just tried to find
110 meter hurdle and then look at a column for reaction times which are always from
roughly 0.1 to 0.18 seconds.} \eds{is it correct now?} \of{Yes}

@@

While it is theoretically possible that athletes got faster 
between 2019 and 2022 or 2022 and 2023, it is important to 
remember that these are some of the fastest sprinters on the planet, and thus
improving on elite times is not easy.  Once again we repeat a similar procedure
as in Section~\ref{sec:datanational}; we assign each athlete to be a cluster and
within each cluster label times to be from either the 2019, 2022, or 2023 World
Championships. When comparing 2019 and 2022 we look at data from 15
athletes, with two clusters of size two, three clusters of size three, five 
clusters of size four, two clusters of size five, and three clusters of size six.
When comparing 2022 and 2023 we look at data from 22 athletes, with seven 
clusters of size two, three clusters of size three, six clusters of size four,
and six clusters of size five.  
%Although 2022 is presented three times in
%Figure~\ref{fig:RankScatterplots}, the subset of athletes differs each time.
%There may be athletes who competed in 2019 and 2022 and had their national
%championships data available, but they did not compete in 2023, and thus their
%times appear only in the top two graphs. \eds{top and bottom graphs?}
\eds{also, you will likely need to make Fig 1 smaller as the caption runs off
the page (even before my edits).}
\of{It should be smaller now}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{RankScatterPlots}
  \caption{The top graph shows the reaction times of athletes who competed at 
	least once each at the 2019 and 2022 World Championships. The middle graph 
	shows the reaction times of athletes who competed at least once each at the 
	2022 and 2023 World Championships. The bottom graph shows the reaction times 
	of athletes who competed at least once at their country's
  2022 national championships and the 2022 World Championships.
	\eds{check x-axis labels in the bottom plot: last three should be 2022?}
  \of{I don't think so. Its a comparison of 2022 national vs international so
  I abbreviated it as Nat. and Inter. is that not ok?}}
  \label{fig:RankScatterplots}
\end{figure}

Figure~\ref{fig:RankScatterplots} shows the data that will be analyzed
with the rank-based methods (described in Section~\ref{sec:rank} with results in 
Section~\ref{subsec:Results_Rank}). In each of the three graphs, the reaction 
times for athletes who competed
in the 2022 finals \eds{World Championships?} are compared to their reaction
times at another at which they competed.  
It is worth noting that each instance of the 2022 data is different, because
those who competed at both the 2019 and 2022 World Championships will be
different athletes than those who competed at both the 2022 and 2023 World
Championships. This data was only for athletes who competed
in at least one race at each championship, however it is possible for athletes
to have competed in one race in 2019 and two races in 2022. 
\eds{did we not already cover the last two sentences in the previous paragraph, 
or are you trying to say something different here?} \of{yes you are correct. I
commented out the other one because I think this one flows better.}  The 2022 World
Championship Heats and Semifinals both had on average lower reaction times than
their respective 2019 races, and the 2022 finals were substantially faster. It
is also worth noting is that the fastest reaction time in the 2022 World
Championships Finals and Semifinals both came from Devon Allen, but the 
difference between 0.101 and 0.099
seconds was the difference in what caused him to be disqualified. 
%\eds{Please reorganize the labels so they are either in one of the following 
%orders:
%2019 Heat, 2019 Semi, 2019 final, 2022 Heat, 2022 Semi, 2022 final  OR
%2019 Heat, 2022 Heat, 2019 Semi, 2022 Semi, 2019 final, 2022 final.  
%(I am not sure which is better, but perhaps the latter?)}
%\of{I re-worked this graph completely, please let me know what you think.  For
%each graph I currently list them in chronological order: so 2022 then 2023 but
%I list 2022 second when comparing it to 2019/rank comparison.  Please let me
%know your thoughts on this and if I should list 2022 and then 2023 for visual
%consistency.}
%\eds{I think the graphs look good}

We also started to perform a similar analysis to compare the times of female 
athletes who competed at the 2019 and 2022 World Track and Field Championships,
however as our results were not significant, we decided to incorporate another
method to investigate the times at the 2022 World Championships and thus we
employ the GLMM, which looks at historical men's data to try to determine
whether Allen's time and disqualification are reasonable. 
\eds{I would suggest removing this paragraph, particularly if you still include
the comments about female reaction times in the Concluding Remarks. 
No analysis has been done yet, nor have the methods been described yet, so it 
does not make sense in my opinion 
to refer to significance or another analysis method here.}
\of{The reason I added it was because Prof Yan recommended switching the
order of glmm and rank and talking about the GLMM partly because the women's
data was not significant. I'm not sure how else to say this or if it needs to
be said at all.  I agree that it doesn't make a ton of sense to talk about the 
results for the women's data so early.  Prof Yan, can you please include your
input? Maybe we keep the rank before glmm but don't bring up women's data until
the end?}

@@

Data was taken from the World Athletics 
and covers the men's 110 meter hurdles from 1999 to 2023.
Although women's data is available, we chose not to include women's data in the
analyses to follow due to the numerous papers citing differences in men and
women's reaction times \citep[e.g.,][]{lipps2011implications, 
babicc2009reaction, panoutsakopoulos2020gender}.
\eds{you should mention the data source earlier, possible in Section 2 
before Section 2.1 if it applies to Section 2.1.1, and also the sources of the 
national data for 2.1.2 (i.e., pointing to the info in Appendix).}
\of{I want to emphasize the differences in the data so I wanted to try and
compartmentalize everything that was different. We chose not to include women's
data for the GLMM because of cited papers and chose not to include women's
data for the signed-rank because the results were not highly significant. So if
I want to talk about that in section 2, I feel like I would have to explicitly
justify the exclusion of the women's data which then goes back to the
conversation above. Also Prof Yan recommended I remove the appendix, so now the
sources are in the Data README.}
For the creation of the GLMM model we focus on the reaction times during the 
semifinal and final heats only, as reaction times from preliminary heats are 
often not as fast as those in later heats \citep[e.g.,][]{zhang2021correlation}.

@@

With our small sample size, the asymptotic normal distribution may not be
reliable, so we use 1000 random permutations to simulate the null distribution
of the test statistic.  
\eds{simulate the null distribution of the test statistic?}
\of{Resolved}
%\eds{I do not understand what is meant in the text after the colon in the 
%previous sentence.  Please rephrase.}
%\of{I'm not sure what that was, I removed it now.}
This method is available from the \texttt{clusWilcox.test()} function
with \texttt{method = `ds'} (for \underline{D}atta and \underline{S}atten) and 
\texttt{exact = TRUE} from R package
\texttt{clusrank} \citep{jiang2020wilcoxon}.

@@

Table~\ref{tab:Clusrankresults} shows the results generated from six 
\eds{three?}
\of{Theres now 6 results in the table, if we remove the asymptotic results then
it should be 3}
tests in R.
We use a \texttt{clusWilcox.test()} with \texttt{exact = TRUE} and 
\texttt{B = 1000} indicating that the random permutation size was chosen to be 
1000, as discussed previously in Section~\ref{sec:rank}.  
%As the sample sizes are small, 65 and 49, we use \texttt{exact = TRUE} as
%that is preferred.  
%In both cases, 
The national versus international comparison
produced lower p values than the 2019 versus 2022 comparison, with the
national vs international permutation test producing a p value of 
$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  This
is a highly significant test result and shows that athletes who competed at the
2022 World Track and Field Championships were on average, significantly faster
at reacting than they were a few months prior.  For the 2019 versus 2022
same athlete comparison with permutation argument, the p value was 
$2.9\cdot10^{-3}$, which is significant at $\alpha = 0.01$.  The result of
comparing 2022 and 2023 was very similar to the above result, as the p value
from the permutation test was $3.0\cdot10^{-3}$.  Taken together, these three
tests help show that athletes competing at the 2022 World Track and Field
Championships had faster reaction times than when they had competed previously.

\eds{not sure the \texttt{exact = FALSE} results need to be reported here or in
Table 4 when we say we use \texttt{exact = TRUE} here and in the previous section.}
\of{I am not going to talk about the asymptotic results but will show them in
table.I changed the exact=TRUE and exact = FALSE text to say permutation and 
asymptotic}
these

\eds{need to comment on 2022 vs 2023 results}
\of{Resolved}

@@

\begin{table}
  \centering
  \caption{Estimated parameters for the two effect GLMM on various data sets. SE
  is short for standard error. \eds{data in table needs to be revisited and table
	needs to be referenced within the text. Also, it needs to be clarified 
	somewhere that ``Old Pooled'' excludes 2022 and ``All Pooled'' includes 2022.}
  \of{Resolved}} 
  \begin{tabular}{c c c c c c}
      \toprule
      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $\delta_{i/j}$ \\
      \midrule
      Year Effect Only & $-2023.5$ & N/A & 0.048 & $-1.857$ & 0.152\\
      Heat Effect Only & $-2070.7$ & 0.073 & N/A & $-1.890$ & 0.135\\
      Full model       & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
      \bottomrule
  \end{tabular}
  \label{tab:Gamma_parameters}
\end{table}

@@

This paper does not consider data from the women's 100 meter hurdle, and although
that was initially because of concerns over reaction time differences in men and
women, exploratory analysis showed that 2022 was at best only
slightly faster relative to other championships going back to 2003 
%\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
%the intro, unless he only considered sprint and not hurdles?} \of{Johnson's
%results considered the mens 100m, mens 110m hurdles, womens 100m but not the 
%womens 100m hurdles.} 
Likewise, the rank-based comparisons resulted in p values
that were not significant. 
\eds{could that be because of even smaller sample sizes?}
\of{Perhaps? I rank the pemutation test for the women's data and the
p value was roughly 0.8.}
  The lack of evidence from the women's data makes it 
harder to conclude that the sensors used at the 2022 championships were faulty, 
although it is unknown if the sensors for the men's 110 meter hurdle and women's
100 meter hurdle were the same.  
\eds{is any calibration of the sensors needed (e.g. for weight?) that might
also play a role for women vs men. (My husband asked me about this as a possible 
reason for the difference in results for men and women, but he did not know
if any calibration is actually done.)}
\of{I honestly am not sure, do you think that this is worth investigating, or
beyond the scope of this paper? I am not sure if the mechanics of Seiko's
technology is public knowledge.}

@@

\eds{what happened to the Appendix?}
\of{Prof Yan recommended I remove it}
